---
title: "Week 4 Rundown"
author: "James Rundle, Bradley Odac, Avi"
date: "9/17/2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
## Month End Balance Sheet
library(readxl)
month_end_balances <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Month end balances ", col_types = c("numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))

month_end_balances$mortgage_flag = factor(month_end_balances$mortgage_flag )

daily_interactions_WF <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Daily interactions with WF")

daily_interactions_WF$Des1 = factor(daily_interactions_WF$Des1)
levels(daily_interactions_WF$Des1)


levels(month_end_balances$mortgage_flag)
mean(month_end_balances$age)
summary(month_end_balances)
```


```{r include=FALSE}
library(readxl)

WF_CardData <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Daily use of a WF credit card ")

Cust12 <- data.frame(WF_CardData[WF_CardData$masked_id == 12,])
Cust1 <- data.frame(WF_CardData[WF_CardData$masked_id == 1,])

Cust12_Des1 <- factor(Cust12$Des1)
Cust12_Des2 <- factor(Cust12$Des2)
Cust12_Des3 <- factor(Cust12$Des3)



```




# Types of Purchases for Masked Id
Sorted the purchases made with credit cards by descriptor 2.
We find Retail and Entertainment are top purchases

```{r}
Top_Cat2<- (head(sort(summary(Cust12_Des2),decreasing = TRUE))[1:3])
print(Top_Cat2)

```

# Most descriptive purchases from sorted descriptor 2
Used the top categories, Entertainment and Retail, to show every purchase
Maybe some regex to find similar qualites

```{r}
Entertainment_Cat <- print(subset(Cust12, subset = Des2 == "ENTERTAINMENT")$Des3)

Retail_Cat <- print(subset(Cust12, subset = Des2 == "RETAIL / DEPARTMENT STORES")$Des3)
```


# Purchase Statistics
Here are some descriptive statistics for the customers with Masked ID 1 and 12

We can use this data to see in which quartile a majority of their purchases lie and classify them as a "spending type"

```{r include=FALSE}

PstatsID12 <- list(Stats = summary(Cust12$Payment), Outliers = boxplot(Cust12$Payment)$out)
PstatsID1 <- list(Stats = summary(Cust1$Payment), Outliers = boxplot(Cust1$Payment)$out)
### arranging data by cust id
```

```{r}
BigList <- list(PstatsID1, PstatsID12)
## Had Trouble figuring out how to  append to big list
print(BigList)
```

######
# Using data to predict

Here I am using the data from checking Balance, Savings Balance, ATM withdrawls, and Age to predict
if the person will make purchases about 1000 dollars.

Now Im not really sure if this is what the function is displaying but we're getting something out of it...


```{r}
merged_data <- merge(month_end_balances,WF_CardData,by="masked_id")
library(randomForest)
fit <- randomForest(as.factor(Payment> 1000) ~ check_bal_altered +
                      sav_bal_altered + atm_withdrawls_cnt + age,
                      data= merged_data, 
                      importance=TRUE, 
                      ntree=2000)
print(fit)

```

# Variable importance Plot
Almost all values do not seem to affect accuracy but savings balance, check balance, and age all seem to  decrese the GINI sharply. If i interpret GINI correctly that means that these variables have much more relevance to the event of a Payment being greater than 1000.


```{r}
varImpPlot(fit)

```