---
title: "Week 4 Rundown"
author: "James Rundle, Bradley Odac, Avi"
date: "9/17/2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
## Month End Balance Sheet
library(readxl)
month_end_balances <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Month end balances ", col_types = c("numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))

month_end_balances$mortgage_flag = factor(month_end_balances$mortgage_flag )

daily_interactions_WF <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Daily interactions with WF")

daily_interactions_WF$Des1 = factor(daily_interactions_WF$Des1)
levels(daily_interactions_WF$Des1)


```


```{r include=FALSE}
library(readxl)

WF_CardData <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Daily use of a WF credit card ")


Cust12 <- data.frame(WF_CardData[WF_CardData$masked_id == 12,])
Cust1 <- data.frame(WF_CardData[WF_CardData$masked_id == 1,])

Cust12_Des1 <- factor(Cust12$Des1)
Cust12_Des2 <- factor(Cust12$Des2)
Cust12_Des3 <- factor(Cust12$Des3)

summary(factor(WF_CardData$Des2))

Popset <- (subset(WF_CardData, subset = Des2 == "OTHER")$Des3)

CardData <- read_excel("~/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Daily use of a WF credit card ")
CardData['MiniDes2'] <- 0

```

```{r}
##Hope this works to salvage Data
#CardData <- WF_CardData2
bills <- list("INSURANCE / FINANCIAL SERVICES","TAX","UTILITIES / TELECOM")
entertainment <- list("ENTERTAINMENT","RESTAURANTS")
personal<- list("PET / VETERINARY","PERSONAL SERVICES","EDUCATION","HEALTHCARE / PHARMACY","LODGING")
transport<- list("AIRLINES / TRANSPORTATION","AUTO / GAS")
work<- list("BUILDING SUPPLY / WHOLESALE","OFFICE SUPPLY / STATIONERY","HOUSEHOLD","CONTRACTOR / BUSINESS SERVICES","POSTAGE / DELIVERY")
common<- list("GROCERIES","RETAIL / DEPARTMENT STORES")
charity <- list("CHARITY / COMMUNITY ORG.")
other <- list("OTHER", NA)


TrimmedDesList <- list("BILLS"= bills,"ENTERTAINMENT"=entertainment,"PERSONAL CARE"=personal,"TRANSPORTATION"=transport,"WORK RELATED"=work,"COMMON PURCHASES"=common,"CHARITY"=charity,"OTHER"=other)

```








```{r include=FALSE}

PstatsID12 <- list(Stats = summary(Cust12$Payment), Outliers = boxplot(Cust12$Payment)$out)
PstatsID1 <- list(Stats = summary(Cust1$Payment), Outliers = boxplot(Cust1$Payment)$out)
### arranging data by cust id
##50 different masked ID
```










### Condensing Des2 Catagories down to 7
Running randomForests last week I felt like there were too many catagories to accurately use.
Des 2 had the least descriptors coming in at 28. I felt it was a good middle ground to classify each "purchase" to a condensed list of 7 catacories.
After wasting way too much time trying to use for loops to propagate a new column I just bruteforced it using nested lists of the 28 descriptors in the Des2 column

```{r}
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$BILLS] <- "BILLS"
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$ENTERTAINMENT] <- "ENTERTAINMENT"
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$`PERSONAL CARE`] <- "PERSONAL CARE"
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$TRANSPORTATION] <- "TRANSPORTATION" 
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$`WORK RELATED`] <- "WORK RELATED" 
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$`COMMON PURCHASES`] <- "COMMON PURCHASES"
CardData$MiniDes2[CardData$Des2 %in% TrimmedDesList$CHARITY] <- "CHARITY" 
CardData$MiniDes2[CardData$Des2 == "OTHER" ] <- "OTHER" 
CardData$MiniDes2[ is.na(CardData$Des2)] <- "OTHER" 

```
## Using the new MiniDes2 column to run randomForests

Well...It's different. I think Im going in the right direction but need some work. Right now the catagorical Random forest appears to only predict "COMMON PURCHASES". Thats not super great.

The regression forest is telling me 
*"The response has five or fewer unique values.  Are you sure you want to do regression?"*
Still coming up with no culprits after some intense googling.

```{r}
MergedWMini <- merge.data.frame(month_end_balances,CardData,by="masked_id", all.x = TRUE )

MergedWMini$MiniDes2 = as.factor(MergedWMini$MiniDes2)

library(randomForest)
CatPredict <- randomForest(as.factor(MiniDes2) ~  checking_acct_ct + online_bank_cnt + direct_phone_cnt + atm_withdrawls_cnt +  age  + mobile_bank_cnt + check_bal_altered + sav_bal_altered,
                      data=MergedWMini, mtry = 3 ,localImp = TRUE,
                      importance=TRUE,na.action = na.omit, 
                      ntree=2000)
print(CatPredict)
varImpPlot(CatPredict)




PayRegr <- randomForest((Payment > 1000) ~ MiniDes2 + check_bal_altered + sav_bal_altered +atm_withdrawls_cnt +  age,
                      data=MergedWMini, mtry = 2 ,localImp = TRUE,
                      importance=TRUE,na.action = na.omit, 
                      ntree=2000)
print(PayRegr)
varImpPlot(PayRegr)

```

### trying to make mindes2 applicable for regression
HALP
```{r}
OrdMiniDes <- MergedWMini$MiniDes2
OrdMiniDes <- as.numeric(OrdMiniDes)-7
```
# Purchase Statistics
Here created dataframe collecting data on each of the 50 masked ids. The top two "purchase types" for each customer.

As we can see some accounts do not have credit cards, but most are "Common purchases" and "Entertainment"/ "Transportation". I think this is a good start but the classifying needs to be tweaked to supply some really valuable/workable information. Suggestions?

Each masked id also contains the descriptive statistics for the Payments(charges) on their creditcard. I would like to find out a percentage of purchases made within each  range. If most of their purchases are high dollar, they need to be marketed towards differently than someone who a majority of their purchases are <50 from grocery stores or something.
```{r}
### Fill dataframe with #1 des, #2 des, descriptive stats,

CardStats <- as.data.frame(matrix(ncol = 9, nrow = 50))
colnames(CardStats) <- c('masked_id',"1stDes",'2ndDes','MinofCharge','Q1ofCharge','MedofCharge','Q3ofCharge','MaxofCharge','MeanofCharge')

for(i in 1:50){
    Top2 <- summary(as.factor(MergedWMini[MergedWMini$masked_id == i,"MiniDes2"]))
    Top2 <- sort(Top2, decreasing = TRUE)
    CardStats$masked_id[i] <- i
    CardStats$'1stDes'[i] <- names(Top2[1])
    CardStats$'2ndDes'[i] <- names(Top2[2])
    CardStats$MinofCharge[i]<- summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['Min.']
    CardStats$Q1ofCharge[i]<-summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['1st Qu.']
    CardStats$MedofCharge[i]<-summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['Median']
    CardStats$Q3ofCharge[i]<-summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['3rd Qu.']
    CardStats$MaxofCharge[i]<-summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['Max.']
    CardStats$MeanofCharge[i]<-summary((MergedWMini[MergedWMini$masked_id == i,"Payment"]))['Mean']

}

print(head(CardStats,10))

```
